---
title: "Virulence Evolution in Little Appleton"
author: "Meghan Duffy"
date: "4/2/2021"
output: html_document
---

Data and code for paper on evolution during a 2017 epidemic of Pasteuria ramosa in Little Appleton Lake.
Authors:
Camden D. Gowler
Haley Essington
Patrick A. Clay
Bruce O'Brien
Clara L. Shaw
Rebecca Bilich
Meghan A. Duffy

# Analysis of data collected on Little Appleton's 2017 Pasteuria epidemic

```{r, load libraries, include=FALSE}
# Load the libraries & set theme for figure export
library(here)
library(tidyverse)
library(cowplot)
library(lme4)
library(ggtext)
library(emmeans)

theme_set(theme_cowplot(font_size = 14))
```

```{r, load dataload, include = FALSE}
# Tell R where files are stored
here::i_am("scripts/VirulenceEvolution.Rmd")

```

## Field data on infection prevalence and host density

```{r, field infection data}
lilApp <- readr::read_csv(here("data/lilApp.csv")) 

# do this to see which parasites are there (past, micg, spiro)
ttt <- lilApp %>% 
  select(Uninfected.Males:M.Past...Spider) %>%
  gather("Parasite", "Count",Uninfected.Males:M.Past...Spider) %>%
  group_by(Parasite) %>%
  summarise(totinfe = sum(Count)) %>%
  filter(totinfe > 0)

past <- lilApp %>% # find columns with pasteuria in title
  select(contains("past"))
lilApp$pasteuria.inf <- rowSums(past)

micg <- lilApp %>% 
  select(contains("mic"))
lilApp$micg.inf <- rowSums(micg)

spiro <- lilApp %>% 
  select(contains("spiro"))
lilApp$spiro.inf <- rowSums(spiro)

LilApp_allpara <- lilApp %>%
  mutate(pasteuria.prev = pasteuria.inf/Total,
         micg.prev = micg.inf/Total,
         spiro.prev = spiro.inf/Total)

# super contrived way of doing this 
sample_vec <- c(rep(NA, 11))
collect <- c(3, 5, 7)
for (i in 1:length(collect)) {
  x <- collect[i]
  sample_vec[x] <- paste0("collected",i)
}
sample_vec[is.na(sample_vec)] <- "regular"
LilApp_allpara$sampling <- sample_vec
LilApp_allpara$sampling <- as.factor(LilApp_allpara$sampling)

str(LilApp_allpara)
summary(LilApp_allpara)

LA_pastprev_plot <- ggplot() +
  geom_line(data = LilApp_allpara, aes(x = Julian.Day, y = pasteuria.prev*100), size = 1.1, alpha = 0.7) +
  geom_point(data = LilApp_allpara, aes(x = Julian.Day, y = pasteuria.prev*100, fill=sampling), size = 5, alpha = 1.0, shape = 21, show.legend = FALSE) +
  ylim(0, 0.45*100) +
  scale_x_continuous(breaks = c(213, 244, 274, 305), labels = c("Aug-1", "Sept-1","Oct-1", "Nov-1")) +
  scale_fill_manual(values=c("#5445b1", "#749dae", "#f3c483",'#7E7E7E')) +
  xlab("") +
  ylab("Prevalence \n(%)") 
LA_pastprev_plot

LA_density_plot <- ggplot() +
  geom_line(data = LilApp_allpara, aes(x = Julian.Day, y = Density), size = 1.1, alpha = 0.7) +
  geom_point(data = LilApp_allpara, aes(x = Julian.Day, y = Density, fill=sampling), size = 5, alpha = 1.0, shape = 21, show.legend = FALSE) +
  scale_y_log10() +
  scale_x_continuous(breaks = c(213, 244, 274, 305), labels = c("Aug-1", "Sept-1","Oct-1", "Nov-1")) +
  scale_fill_manual(values=c("#5445b1", "#749dae", "#f3c483",'#7E7E7E')) +
  xlab("Date") +
  ylab("*D. dentifera* density \n(per m^2)") +
  theme(axis.title.y = ggtext::element_markdown())
LA_density_plot

fielddataplot <- plot_grid(LA_pastprev_plot, LA_density_plot, labels = "auto", ncol = 1, align = "v")
fielddataplot

ggsave(here("figures", "fielddataplot.jpg"), fielddataplot, units = "in", width = 5, height = 7, dpi = 300)
```
## Infection Assay Data

### Checking out sample sizes
```{r, initial manipulations of infection assay data}
# Load the data for analyzing virulence
LA_infected <- readr::read_csv(here("data/LA_infected.csv")) 

# Get rid of time 4 (which had only one host clone)
LA_infected <- subset(LA_infected, host_time_point != "4")

# Calculating the proportion that never reproduced
LA_infected <- LA_infected %>%
  mutate(anyrepro = if_else(clutches == "0", "0", "1"))

LA_infected$anyrepro <- as.numeric(LA_infected$anyrepro)

# Adding a column for log spores
LA_infected$logspores <- log10(LA_infected$spores_tot)
LA_infected$lnspores <- log(LA_infected$spores_tot)


```

```{r, figuring out number of reps for different clone-parasite combinations}
LA_infected_samplesizes <- LA_infected %>%
  count(clone, para_time_point)

# save
write.csv(LA_infected_samplesizes, here("tables", "LA_infected_samplesizes.csv"), row.names = F)

```

## Infected vs. uninfected comparison

```{r, first getting data in shape for plotting and analysis}
LA_infected$clone <- as.factor(as.character(LA_infected$clone))
LA_infected$host_time_point <- as.factor(as.character(LA_infected$host_time_point))
LA_infected$para_time_point <- as.factor(as.character(LA_infected$para_time_point))

LA_infected$exposed <- ifelse(LA_infected$para_time_point == "none", 'unexposed', 'infected')

# Note: all of the animals in this dataset are either controls or infected, so don't need to worry about exposed but uninfected animals

LA_infected_summary <- LA_infected %>%
  group_by(exposed, host_time_point, clone, para_time_point) %>%
  summarise(meanlifespan = mean(lifespan), meanclutches = mean(clutches)) 

```



```{r, plotting lifespan data for infvcontrol}
lifespaninfvcontrol <- ggplot(LA_infected_summary, 
                               aes(x=exposed, y=meanlifespan)) + 
  geom_violin() + 
  geom_jitter(shape=16, position=position_jitter(width=0.3, height=0), alpha = 0.5, color = '#5c1a33', size = 2) + 
  ylab("Mean lifespan \n(days)")  +
  theme_cowplot() +
  theme(axis.title.x = element_blank())


lifespaninfvcontrol

```


```{r, plotting fecundity data for infected v control comparison}


# making a plot comparing exposed vs. unexposed
fecundityinfvcontrol <- ggplot(LA_infected_summary, 
                               aes(x=exposed, y=meanclutches)) + 
  geom_violin() + 
  geom_jitter(shape=16, position=position_jitter(width=0.3, height=0), alpha = 0.5, color = '#5c1a33', size = 2) +
  ylab("Mean # clutches")  +
  theme_cowplot() +
  theme(axis.title.x = element_blank())

fecundityinfvcontrol
```


Let's arrange the lifespan & fecundity plots into a single figure:
```{r, single plot of inf v unexposed}
infvcontrolplot <- plot_grid(lifespaninfvcontrol, fecundityinfvcontrol, labels = "auto", ncol = 1, align = "v")
infvcontrolplot

ggsave(here("figures", "infvcontrolplot.jpg"), infvcontrolplot, units = "in", width = 5, height = 8, dpi = 300)

```


```{r, analysis of host lifespan and reproduction for inf v control using unpruned data}

infvcontrollifespanmodel<-glmer(lifespan ~ exposed + (1 | clone), family = "poisson", data=LA_infected) 

summary(infvcontrollifespanmodel)

# Testing for overdispersion (using code from Ben Bolker via Michelle Fearon)
overdisp_fun <- function(model) {
  rdf <- df.residual(model)
  rp <- residuals(model,type="pearson")
  Pearson.chisq <- sum(rp^2)
  prat <- Pearson.chisq/rdf
  pval <- pchisq(Pearson.chisq, df=rdf, lower.tail=FALSE)
  c(chisq=Pearson.chisq,ratio=prat,rdf=rdf,p=pval)
}

overdisp_fun(infvcontrollifespanmodel)

#very overdispersed, so let's go to a negative binomial distribution
infvcontrollifespanmodel.nb <-glmer.nb(lifespan ~ exposed + (1 | clone), data=LA_infected) 

overdisp_fun(infvcontrollifespanmodel.nb)

summary(infvcontrollifespanmodel.nb)


infvcontrolrepro<-glmer(clutches ~ exposed + (1 | clone), family = "poisson", data=LA_infected) 

overdisp_fun(infvcontrolrepro)

infvcontrolrepro.nb <-glmer.nb(clutches ~ exposed + (1 | clone), data=LA_infected) 

overdisp_fun(infvcontrolrepro.nb)

summary(infvcontrolrepro.nb)

```


## Infection prevalence in infection assays
```{r, import dataset that will allow analysis of prevalence in infection assays}
# The dataset we were working with above only has infected animals or controls. Let's upload the dataset that will also include animals that were exposed but did not become infected
LA_full <- readr::read_csv(here("data/LA_full.csv")) 

# since we want to look at infectivity, let's get rid of the control/unexposed animals
LA_exposed <- LA_full %>%
  subset(para_time_point != "none")

# get rid of time 4 hosts, since we didn't have parasites from that time point
LA_exposed <- LA_exposed %>%
  subset(host_time_point != '4')

# filter out ones that are missing spore counts
LA_exposed <- LA_exposed %>%
  drop_na(spores_tot)

# Looking for clones that had very few animals in the experiment
LA_exposed_samplesizes <- LA_exposed %>%
  count(clone, para_time_point)

LA_exposed_samplesizes$remove <- ifelse(LA_exposed_samplesizes$n < 3, 'y', 'n')

# save
write.csv(LA_exposed_samplesizes, here("tables", "LA_exposed_samplesizes.csv"), row.names = F)


# Using spore yield to determine if it was infected; this is more sensitive than the 'infection' column, which was based on a quick visual assessment at death -- that inspection missed some that had very low spore densities
LA_exposed$infected <- ifelse(LA_exposed$spores_tot == 0, 0, 1)


# Calculate number infected and total sample sizes
LA_exposed_prevalence <- LA_exposed %>%
  group_by(clone, host_time_point, para_time_point) %>%
  summarise(total = n(), inf = sum(infected)) 
  
# Calculate proportion infected per clone
LA_exposed_prevalence$propinf <- LA_exposed_prevalence$inf/LA_exposed_prevalence$total

LA_exposed_prevalence$host_time_point <- as.character(LA_exposed_prevalence$host_time_point)

infprevplot <- ggplot(LA_exposed_prevalence, 
                     aes(x=host_time_point, y=propinf, fill=para_time_point)) + 
  geom_violin(position=position_dodge(1)) + 
  geom_jitter(shape=16, position=position_jitter(width=0.3, height=0), alpha = 0.5, show.legend = FALSE) +
  scale_fill_manual(values=c("#5445b1", "#749dae", "#f3c483")) + 
  scale_color_manual(values=c("#5445b1", "#749dae", "#f3c483")) +
  xlab("Host time point") +ylab("Proportion infected") +
  labs(color= "Parasite time point") + labs(fill="Parasite time point") +
  theme_cowplot()

infprevplot

infprevplota <- LA_exposed_prevalence %>%
  filter(host_time_point == para_time_point) %>%
  ggplot(aes(x=host_time_point, y=propinf, fill=para_time_point)) + 
  geom_violin(position=position_dodge(1), show.legend = FALSE) + 
  geom_jitter(shape=16, position=position_jitter(width=0.3, height=0), alpha = 0.5, show.legend = FALSE) +
  scale_fill_manual(values=c("#5445b1", "#749dae", "#f3c483")) + 
  scale_color_manual(values=c("#5445b1", "#749dae", "#f3c483")) +
  ylim(0,1.00) +
  xlab("Host time point") +ylab("Proportion infected (contemporary parasites)") +
  labs(color= "Parasite time point") + labs(fill="Parasite time point") +
  theme_cowplot()

infprevplota

infprevplotb <- LA_exposed_prevalence %>%
  filter(host_time_point == 3) %>%
  ggplot(aes(x=para_time_point, y=propinf, fill=para_time_point)) + 
  geom_violin(position=position_dodge(1), show.legend = FALSE) + 
  geom_jitter(shape=16, position=position_jitter(width=0.3, height=0), alpha = 0.5, show.legend = FALSE) +
  scale_fill_manual(values=c("#C0C0C0", "#f3c483")) + 
  scale_color_manual(values=c("#C0C0C0", "#f3c483")) +
  ylim(0,1.00) +
  xlab("Parasite time point") +ylab("Proportion infected (time 3 hosts)") +
  theme_cowplot()

infprevplotb

infprevplot2 <- plot_grid(infprevplota, infprevplotb, labels = "auto", ncol = 2, align = "v")
infprevplot2

ggsave(here("figures", "infprevplot.jpg"), infprevplot2, units = "in", width = 8, height = 5, dpi = 300)
```

```{r, analyze infection prevalence data}
LA_exposed_prevalence_sametime <- subset(LA_exposed_prevalence, host_time_point == para_time_point)

prevmodel1<-glmer(cbind(inf, total-inf) ~ para_time_point + (1 | clone), family='binomial', data=LA_exposed_prevalence_sametime) 

# Testing for overdispersion (using code from Ben Bolker via Michelle Fearon)
overdisp_fun <- function(model) {
  rdf <- df.residual(model)
  rp <- residuals(model,type="pearson")
  Pearson.chisq <- sum(rp^2)
  prat <- Pearson.chisq/rdf
  pval <- pchisq(Pearson.chisq, df=rdf, lower.tail=FALSE)
  c(chisq=Pearson.chisq,ratio=prat,rdf=rdf,p=pval)
}

overdisp_fun(prevmodel1)

summary(prevmodel1)

LA_exposed_prevalence_sametime$para_time_point <- as.factor(LA_exposed_prevalence_sametime$para_time_point)

# Testing for overall effect of parasite time point
drop1(prevmodel1,test="Chisq")

# Now comparing the two sets of spores that were used to infect time 3 hosts
LA_exposed_prevalence_time3 <- subset(LA_exposed_prevalence, host_time_point == '3')

prevmodel2<-glmer(cbind(inf, total-inf) ~ para_time_point + (1 | clone), family='binomial', data=LA_exposed_prevalence_time3) 

overdisp_fun(prevmodel2)

summary(prevmodel2)

# Getting chi-square to be parallel to initial analysis
drop1(prevmodel2,test="Chisq")
```


## Contemporary virulence
Comparing hosts exposed to parasites from the same time point

```{r, get dataset for contemporary exposures}

LA_sametime <- LA_infected %>%
  filter(para_time_point != "none") 

LA_sametime$para_time_point <- as.factor(LA_sametime$para_time_point)
LA_sametime$host_time_point <- as.factor(as.character(LA_sametime$host_time_point))

LA_sametime$para_time_point <- droplevels(LA_sametime$para_time_point)

LA_sametime <- subset(LA_sametime, host_time_point == para_time_point)
```

```{r, lifespan for hosts exposed to contemporary parasites}
LA_sametime$paragrowth <- (LA_sametime$spores_tot)/(LA_sametime$lifespan)


LA_sametime_summary <- LA_sametime %>%
  group_by(clone, para_time_point) %>%
  summarise(meanlifespan = mean(lifespan), meanclutches = mean(clutches), meanspores = mean(spores_tot), meanlogspores = mean(logspores), meanparagrowth = mean(paragrowth)) 


contemplifespanplot <- ggplot(LA_sametime_summary, 
                     aes(x=para_time_point, y=meanlifespan, fill=para_time_point)) + 
  geom_violin(show.legend = FALSE) + 
  geom_jitter(shape=16, position=position_jitter(width=0.3, height=0, seed = 1), alpha = 0.5, show.legend = FALSE) +
  scale_fill_manual(values=c("#5445b1", "#749dae", "#f3c483")) + 
  scale_color_manual(values=c("#5445b1", "#749dae", "#f3c483")) +
  xlab("Parasite time point") +ylab("Mean lifespan \n(days)") +
  labs(color= "Parasite time point") + labs(fill="Parasite time point") +
  theme_cowplot()


contemplifespanplot
```


```{r, analysis of contemporary lifespan}
contemporarylifespan <-glmer(lifespan ~ host_time_point + (1 | clone), family = "poisson", data=LA_sametime) 

overdisp_fun(contemporarylifespan)

summary(contemporarylifespan)

contemporarylifespan.nb <-glmer.nb(lifespan ~ host_time_point + (1 | clone), data=LA_sametime) 

overdisp_fun(contemporarylifespan.nb)

summary(contemporarylifespan.nb)
drop1(contemporarylifespan.nb,test="Chisq")

emmeans(contemporarylifespan.nb, specs = pairwise ~ host_time_point)

```

```{r, reproduction of hosts exposed to contemporary parasites}

contempreproplot <- ggplot(LA_sametime_summary, 
                     aes(x=para_time_point, y=meanclutches, fill=para_time_point)) + 
  geom_violin(show.legend = FALSE) + 
  geom_jitter(shape=16, position=position_jitter(width=0.3, height=0, seed = 1), alpha = 0.5, show.legend = FALSE) +
  scale_fill_manual(values=c("#5445b1", "#749dae", "#f3c483")) + 
  scale_color_manual(values=c("#5445b1", "#749dae", "#f3c483")) +
  xlab("Parasite time point") +ylab("Mean number of clutches") +
  labs(color= "Parasite time point") + labs(fill="Parasite time point") +
  theme_cowplot()


contempreproplot
```

```{r, analysis of contemporary reproduction}
contemporaryclutches <-glmer(clutches ~ host_time_point + (1 | clone), family = "poisson", data=LA_sametime) 

overdisp_fun(contemporaryclutches)

summary(contemporaryclutches)

contemporaryclutches.nb <-glmer.nb(clutches ~ host_time_point + (1 | clone), data=LA_sametime) 

overdisp_fun(contemporaryclutches.nb)

summary(contemporaryclutches.nb)
drop1(contemporaryclutches.nb,test="Chisq")

emmeans(contemporaryclutches.nb, specs = pairwise ~ host_time_point)

```


```{r, spores produced by hosts exposed to contemporary parasites}

contempsporesplot <- ggplot(LA_sametime_summary, 
                     aes(x=para_time_point, y=meanspores, fill=para_time_point)) + 
  geom_violin(show.legend = FALSE) + 
  geom_jitter(shape=16, position=position_jitter(width=0.3, height=0, seed = 1), alpha = 0.5, show.legend = FALSE) +
  scale_fill_manual(values=c("#5445b1", "#749dae", "#f3c483")) + 
  scale_color_manual(values=c("#5445b1", "#749dae", "#f3c483")) +
  xlab("Parasite time point") +ylab("Mean number of \nspores per infected host") +
  labs(color= "Parasite time point") + labs(fill="Parasite time point") +
  scale_y_log10() +
  theme_cowplot()


contempsporesplot
```

```{r, analysis of contemporary spores}
contemporarylogspores <- lmer(logspores ~ host_time_point + (1 | clone), data=LA_sametime) 

overdisp_fun(contemporarylogspores)

summary(contemporarylogspores)

drop1(contemporarylogspores,test="Chisq")

emmeans(contemporarylogspores, specs = pairwise ~ host_time_point)

#R3 asked about the non-log-transformed data, so let's do that
contemporaryspores <- lmer(spores_tot ~ host_time_point + (1 | clone), data=LA_sametime) 

overdisp_fun(contemporaryspores)

summary(contemporaryspores)

drop1(contemporaryspores,test="Chisq")

```


```{r, parasite growth rate for hosts exposed to contemporary parasites}
contempparagrowthplot <- ggplot(LA_sametime_summary, 
                     aes(x=para_time_point, y=meanparagrowth, fill=para_time_point)) + 
  geom_violin(show.legend = FALSE) + 
  geom_jitter(shape=16, position=position_jitter(width=0.3, height=0, seed = 1), alpha = 0.5, show.legend = FALSE) +
  scale_fill_manual(values=c("#5445b1", "#749dae", "#f3c483")) + 
  scale_color_manual(values=c("#5445b1", "#749dae", "#f3c483")) +
  xlab("Parasite time point") +ylab("Parasite growth rate \n(spores per day)") +
  labs(color= "Parasite time point") + labs(fill="Parasite time point") +
  theme_cowplot()


contempparagrowthplot

```


```{r, analysis of contemporary growthrate}

contemporaryparasitegrowth <- lmer(paragrowth ~ host_time_point + (1 | clone), data=LA_sametime) 

overdisp_fun(contemporaryparasitegrowth)

summary(contemporaryparasitegrowth)

drop1(contemporaryparasitegrowth,test="Chisq")

emmeans(contemporaryparasitegrowth, specs = pairwise ~ host_time_point)

```




## Parasite evolution comparison
Comparing hosts from time 3 that were exposed to parasites from time 1 vs. time 3

```{r, get dataset for parasite evolution comparison}
LA_paraevol <- LA_infected %>%
  filter(host_time_point == "3") %>%
  filter(para_time_point != "none")

LA_paraevol$paragrowth <- (LA_paraevol$spores_tot)/(LA_paraevol$lifespan)


LA_paraevol_summary <- LA_paraevol %>%
  group_by(clone, para_time_point) %>%
  summarise(meanlifespan = mean(lifespan), meanclutches = mean(clutches), meanspores = mean(spores_tot), meanlogspores = mean(logspores), meanparagrowth = mean(paragrowth)) 
```


### Lifespan
```{r, lifespan for hosts from time 3}
lifespanparaevol <- ggplot(LA_paraevol_summary, 
                     aes(x=para_time_point, y=meanlifespan, fill=para_time_point)) + 
  geom_violin(show.legend = FALSE) + 
  geom_jitter(shape=16, position=position_jitter(width=0.3, height=0, seed = 1), alpha = 0.5, show.legend = FALSE) +
  scale_fill_manual(values=c("#C0C0C0", "#f3c483")) + 
  scale_color_manual(values=c("#C0C0C0", "#f3c483")) +
  xlab("Parasite time point") +ylab("Mean lifespan \n(days)") +
  labs(color= "Parasite time point") + labs(fill="Parasite time point") +
  theme_cowplot()


lifespanparaevol

```

```{r, analysis of host lifespan for parasite evolution comparison}
paraevollifespan <-glmer(lifespan ~ para_time_point + (1 | clone), family = "poisson", data=LA_paraevol) 

overdisp_fun(paraevollifespan)

summary(paraevollifespan)

paraevollifespan.nb <-glmer.nb(lifespan ~ para_time_point + (1 | clone), data=LA_paraevol) 

overdisp_fun(paraevollifespan.nb)

summary(paraevollifespan.nb)
```


### Reproduction
```{r, reproduction for hosts from time 3}
reproparaevol <- ggplot(LA_paraevol_summary, 
                     aes(x=para_time_point, y=meanclutches, fill=para_time_point)) + 
  geom_violin(show.legend = FALSE) + 
  geom_jitter(shape=16, position=position_jitter(width=0.3, height=0, seed = 1), alpha = 0.5, show.legend = FALSE) +
  scale_fill_manual(values=c("#C0C0C0", "#f3c483")) + 
  scale_color_manual(values=c("#C0C0C0", "#f3c483")) +
  xlab("Parasite time point") +ylab("Mean number of clutches") +
  labs(color= "Parasite time point") + labs(fill="Parasite time point") +
  theme_cowplot()


reproparaevol

```

```{r, analysis of host reproduction for parasite evolution comparison}
paraevolclutches <-glmer(clutches ~ para_time_point + (1 | clone), family = "poisson", data=LA_paraevol) 

overdisp_fun(paraevolclutches)

summary(paraevolclutches)

paraevolclutches.nb <-glmer.nb(clutches ~ para_time_point + (1 | clone), data=LA_paraevol) 

overdisp_fun(paraevolclutches.nb)

summary(paraevolclutches.nb)
```


### Spores
```{r, spore yield for hosts from time 3}
sporesparaevol <- ggplot(LA_paraevol_summary, 
                     aes(x=para_time_point, y=meanspores, fill=para_time_point)) + 
  geom_violin(show.legend = FALSE) + 
  geom_jitter(shape=16, position=position_jitter(width=0.3, height=0, seed = 1), alpha = 0.5, show.legend = FALSE) +
  scale_fill_manual(values=c("#C0C0C0", "#f3c483")) + 
  scale_color_manual(values=c("#C0C0C0", "#f3c483")) +
  xlab("Parasite time point") +ylab("Mean number of \nspores per infected host") +
  labs(color= "Parasite time point") + labs(fill="Parasite time point") +
  scale_y_log10() +
  theme_cowplot()


sporesparaevol

```

```{r, analysis of spore yield for parasite evolution comparison}
paraevollogspores <- lmer(logspores ~ para_time_point + (1 | clone), data=LA_paraevol) 

overdisp_fun(paraevollogspores)

summary(paraevollogspores)

drop1(paraevollogspores,test="Chisq")


#R3 asked about the analysis if we don't log transform the data
paraevolspores <- lmer(spores_tot ~ para_time_point + (1 | clone), data=LA_paraevol) 

overdisp_fun(paraevolspores)

summary(paraevolspores)

drop1(paraevolspores,test="Chisq")
```

### Parasite growth rate
```{r, parasite growth rate for hosts from time 3}
paragrowthparaevol <- ggplot(LA_paraevol_summary, 
                     aes(x=para_time_point, y=meanparagrowth, fill=para_time_point)) + 
  geom_violin(show.legend = FALSE) + 
  geom_jitter(shape=16, position=position_jitter(width=0.3, height=0, seed = 1), alpha = 0.5, show.legend = FALSE) +
  scale_fill_manual(values=c("#C0C0C0", "#f3c483")) + 
  scale_color_manual(values=c("#C0C0C0", "#f3c483")) +
  xlab("Parasite time point") +ylab("Parasite growth rate \n(spores per day)") +
  labs(color= "Parasite time point") + labs(fill="Parasite time point") +
  theme_cowplot()


paragrowthparaevol

```

```{r, analysis of parasite growth rate for parasite evolution comparison}
paraevolparasitegrowth <- lmer(paragrowth ~ para_time_point + (1 | clone), data=LA_paraevol) 

summary(paraevolparasitegrowth)

drop1(paraevolparasitegrowth,test="Chisq")

```

### Combined figure 
```{r, figure with 8 plots}
eightpanelplot <- plot_grid(contemplifespanplot, lifespanparaevol, contempreproplot, reproparaevol, contempsporesplot, sporesparaevol, contempparagrowthplot, paragrowthparaevol, labels = "auto", ncol = 2, align = "v")
eightpanelplot

# now add the title
title1 <- ggplot() +                      # Draw ggplot2 plot with text only
  annotate("text",
           x = 0,
           y = 1,
           size = 6,
           label = "Contemporary hosts",
#           hjust = 0.64,
            hjust = 0.5,
           fontface = "bold") + 
  theme_void()

title2 <- ggplot() +                      # Draw ggplot2 plot with text only
  annotate("text",
           x = 0,
           y = 1,
           size = 6,
           label = "Time 3 hosts",
           hjust = 0.5,
           fontface = "bold") + 
  theme_void()

titles <- plot_grid(title1, title2, nrow = 1,   hjust = 0)

titles

eightpanelplot_withtitles <- plot_grid(
  titles, eightpanelplot,
  ncol = 1,
  # rel_heights values control vertical title margins
  rel_heights = c(0.05, 1)
)

eightpanelplot_withtitles


ggsave(here("figures", "eightpanelplot_withtitles.jpg"), eightpanelplot_withtitles, units = "in", width = 8, height = 12, dpi = 300)
```



# Exposed but uninfected animals

Reviewer 3 was asking about exposed but uninfected animals. Let's make a new dataframe to look at that.
```{r, creating dataframe with exposed but uninfected animals}

# get rid of time 4 hosts from the full dataset, since we didn't have parasites from that time point
LA_full <- LA_full %>%
  subset(host_time_point != '4')

# filter out ones that are missing spore counts
LA_full <- LA_full %>%
  drop_na(spores_tot)

# Looking for clones that had very few animals in the experiment
LA_full_samplesizes <- LA_full %>%
  count(clone, para_time_point)

# save
write.csv(LA_full_samplesizes, here("tables", "LA_full_samplesizes.csv"), row.names = F)


# Using spore yield to determine if it was infected; this is more sensitive than the 'infection' column, which was based on a quick visual assessment at death -- that inspection missed some that had very low spore densities
LA_full$infected <- ifelse(LA_full$spores_tot == 0, 0, 1)

LA_full$exposureclass <- ifelse(LA_full$para_time_point == "none", 'control',
                                ifelse(LA_full$spores_tot > 0, 'infected', 'exposed'))

LA_full_summary <- LA_full %>%
  group_by(exposureclass, host_time_point, clone, para_time_point) %>%
  summarise(meanlifespan = mean(lifespan), meanclutches = mean(clutches)) 

```


```{r, plotting lifespan data for three exposure classes}


lifespan_exposureclass <- ggplot(LA_full_summary, 
                               aes(x=exposureclass, y=meanlifespan)) + 
  geom_violin() + 
  geom_jitter(shape=16, position=position_jitter(width=0.3, height=0), alpha = 0.5, color = '#5c1a33', size = 2) + 
  ylab("Mean lifespan \n(days)")  +
  theme_cowplot() +
  theme(axis.title.x = element_blank())


lifespan_exposureclass

```


```{r, plotting fecundity data for three exposure classes}


# making a plot comparing exposed vs. unexposed
repro_exposureclass <- ggplot(LA_full_summary, 
                               aes(x=exposureclass, y=meanclutches)) + 
  geom_violin() + 
  geom_jitter(shape=16, position=position_jitter(width=0.3, height=0), alpha = 0.5, color = '#5c1a33', size = 2) +
  ylab("Mean number of clutches")  +
  theme_cowplot() +
  theme(axis.title.x = element_blank())

repro_exposureclass
```


Let's arrange the lifespan & fecundity plots into a single figure:
```{r, single plot of three exposure classes}
exposureclassplot <- plot_grid(lifespan_exposureclass, repro_exposureclass, labels = "auto", ncol = 1, align = "v")
exposureclassplot

ggsave(here("figures", "exposureclassplot.jpg"), exposureclassplot, units = "in", width = 5, height = 8, dpi = 300)

```


```{r, analysis of host lifespan and reproduction for three exposure classes}

lifespanmodel_3classes.nb <-glmer.nb(lifespan ~ exposureclass + (1 | clone), data=LA_full) 

overdisp_fun(lifespanmodel_3classes.nb)

summary(lifespanmodel_3classes.nb)

emmeans(lifespanmodel_3classes.nb, specs = pairwise ~ exposureclass)


repromodel_3classes.nb <-glmer.nb(clutches ~ exposureclass + (1 | clone), data=LA_full) 

summary(repromodel_3classes.nb)

emmeans(repromodel_3classes.nb, specs = pairwise ~ exposureclass)

```
