---
title: "Virulence Evolution in Little Appleton"
author: "Meghan Duffy"
date: "4/2/2021"
output: html_document
---

Data and code for paper on evolution during a 2017 epidemic of Pasteuria ramosa in Little Appleton Lake.
Authors:
Camden D. Gowler
Haley Essington
Patrick A. Clay
Bruce O'Brien
Clara L. Shaw
Rebecca Bilich
Meghan A. Duffy

## Analysis of data collected on Little Appleton's 2017 Pasteuria epidemic

```{r, load libraries, include=FALSE}
# Load the libraries & set theme for figure export
library(here)
library(tidyverse)
library(cowplot)
library(lme4)
library(ggtext)

theme_set(theme_cowplot(font_size = 14))
```

```{r, load dataload, include = FALSE}
# Tell R where files are stored
here::i_am("scripts/VirulenceEvolution.Rmd")

# Load the data for analyzing virulence
LA_infected <- readr::read_csv(here("data/LA_infected.csv")) 
```

```{r, field infection data}
lilApp <- readr::read_csv(here("data/lilApp.csv")) 

# do this to see which parasites are there (past, micg, spiro)
ttt <- lilApp %>% 
  select(Uninfected.Males:M.Past...Spider) %>%
  gather("Parasite", "Count",Uninfected.Males:M.Past...Spider) %>%
  group_by(Parasite) %>%
  summarise(totinfe = sum(Count)) %>%
  filter(totinfe > 0)

past <- lilApp %>% # find columns with pasteuria in title
  select(contains("past"))
lilApp$pasteuria.inf <- rowSums(past)

micg <- lilApp %>% 
  select(contains("mic"))
lilApp$micg.inf <- rowSums(micg)

spiro <- lilApp %>% 
  select(contains("spiro"))
lilApp$spiro.inf <- rowSums(spiro)

LilApp_allpara <- lilApp %>%
  mutate(pasteuria.prev = pasteuria.inf/Total,
         micg.prev = micg.inf/Total,
         spiro.prev = spiro.inf/Total)

# super contrived way of doing this 
sample_vec <- c(rep(NA, 11))
collect <- c(3, 5, 7)
for (i in 1:length(collect)) {
  x <- collect[i]
  sample_vec[x] <- paste0("collected",i)
}
sample_vec[is.na(sample_vec)] <- "regular"
LilApp_allpara$sampling <- sample_vec
LilApp_allpara$sampling <- as.factor(LilApp_allpara$sampling)

str(LilApp_allpara)
summary(LilApp_allpara)

LA_pastprev_plot <- ggplot() +
  geom_line(data = LilApp_allpara, aes(x = Julian.Day, y = pasteuria.prev*100), size = 1.1, alpha = 0.7) +
  geom_point(data = LilApp_allpara, aes(x = Julian.Day, y = pasteuria.prev*100, fill=sampling), size = 5, alpha = 1.0, shape = 21, show.legend = FALSE) +
  ylim(0, 0.45*100) +
  scale_x_continuous(breaks = c(213, 244, 274, 305), labels = c("Aug-1", "Sept-1","Oct-1", "Nov-1")) +
  scale_fill_manual(values=c("#5445b1", "#749dae", "#f3c483",'#7E7E7E')) +
  xlab("") +
  ylab("Prevalence \n(%)") 
LA_pastprev_plot

LA_density_plot <- ggplot() +
  geom_line(data = LilApp_allpara, aes(x = Julian.Day, y = Density), size = 1.1, alpha = 0.7) +
  geom_point(data = LilApp_allpara, aes(x = Julian.Day, y = Density, fill=sampling), size = 5, alpha = 1.0, shape = 21, show.legend = FALSE) +
  scale_y_log10() +
  scale_x_continuous(breaks = c(213, 244, 274, 305), labels = c("Aug-1", "Sept-1","Oct-1", "Nov-1")) +
  scale_fill_manual(values=c("#5445b1", "#749dae", "#f3c483",'#7E7E7E')) +
  xlab("Date") +
  ylab("*D. dentifera* density \n(per m^2)") +
  theme(axis.title.y = ggtext::element_markdown())
LA_density_plot

fielddataplot <- plot_grid(LA_pastprev_plot, LA_density_plot, labels = "auto", ncol = 1, align = "v")
fielddataplot

ggsave(here("figures", "fielddataplot.jpg"), fielddataplot, units = "in", width = 5, height = 7, dpi = 300)
```

```{r, dealing with clone-parasite combinations with few reps}
LA_infected_samplesizes <- LA_infected %>%
  count(clone, para_time_point)

# save
write.csv(LA_infected_samplesizes, here("tables", "LA_infected_samplesizes.csv"), row.names = F)

# remove clone-parasite treatments that had fewer than 3 individuals, based on the output of the table above
LA_infected$remove <- ifelse(LA_infected$clone == '17' & LA_infected$para_time_point == "1" | 
           LA_infected$clone == '31' |
           LA_infected$clone == '224' |
           LA_infected$clone == '241' |
           LA_infected$clone == '312' & LA_infected$para_time_point == "none" |
           LA_infected$clone == '312' & LA_infected$para_time_point == "1" |
           LA_infected$clone == '329' & LA_infected$para_time_point == "1" |
           LA_infected$clone == '335' |
           LA_infected$clone == '337' & LA_infected$para_time_point == "1" |
           LA_infected$clone == '343' & LA_infected$para_time_point == "1", 1, 0) 

LA_infected_pruned <- LA_infected %>%
  subset(remove == 0)

# Get rid of time 4
LA_infected_pruned <- subset(LA_infected_pruned, host_time_point != "4")

LA_infected_pruned_samplesizes <- LA_infected_pruned %>%
  filter(para_time_point != "none") %>%
  count(clone, para_time_point)
```

```{r, basic data manipulation}
# Calculating the proportion that never reproduced
LA_infected_pruned <- LA_infected_pruned %>%
  mutate(anyrepro = if_else(clutches == "0", "0", "1"))

LA_infected_pruned$anyrepro <- as.numeric(LA_infected_pruned$anyrepro)

# Adding a column for log spores
LA_infected_pruned$logspores <- log10(LA_infected_pruned$spores_tot)
LA_infected_pruned$lnspores <- log(LA_infected_pruned$spores_tot)
```


```{r, calculating mean host lifespan}
str(LA_infected_pruned)
LA_infected_pruned$clone <- as.factor(as.character(LA_infected_pruned$clone))

# get mean lifespan for each clone*treatment
stats_lifesp <- LA_infected_pruned %>%
  group_by(host_time_point, para_time_point, clone) %>%
  summarise(mn_lifespan = mean(lifespan)) 

# for selecting contemporaneous pairs of hosts/parasites
stats_lifesp$host_time_point <- as.character(stats_lifesp$host_time_point)
stats_lifesp$para_time_point <- as.character(stats_lifesp$para_time_point)

LA_inf <- stats_lifesp %>% # select non-controls
  filter(para_time_point != "none") 

LA_control <- stats_lifesp %>% # select controls
  filter(host_time_point != "4") %>%
  filter(para_time_point == "none") %>%
  ungroup() %>%
  select(-para_time_point)

colnames(LA_control)[3] <- "control_lifespan"

# rejoin infected and control dfs
LA_rel_vir_lifespan <- full_join(LA_inf, LA_control)
LA_rel_vir_lifespan$host_time_point <- as.factor(LA_rel_vir_lifespan$host_time_point)

# make a separate file to use later for infected v. control comparison
LA_rel_vir_lifespan_infvcontrol <- full_join(LA_inf, LA_control)
LA_rel_vir_lifespan_infvcontrol$host_time_point <- as.factor(LA_rel_vir_lifespan_infvcontrol$host_time_point)

LA_rel_vir_lifespan <- as.data.frame(LA_rel_vir_lifespan)

str(LA_rel_vir_lifespan)
```

```{r, calculating mean fecundity}
str(LA_infected_pruned)

# get mean number clutches for each clone*treatment
stats_clutches <- LA_infected_pruned %>%
  group_by(host_time_point, para_time_point, clone) %>%
  summarise(mn_clutches = mean(clutches)) 

# for selecting contemporaneous pairs of hosts/parasites
stats_clutches$host_time_point <- as.character(stats_clutches$host_time_point)
stats_clutches$para_time_point <- as.character(stats_clutches$para_time_point)

LA_inf <- stats_clutches %>% # select non-controls
  filter(para_time_point != "none") 

LA_control <- stats_clutches %>% # select controls
  filter(host_time_point != "4") %>%
  filter(para_time_point == "none") %>%
  ungroup() %>%
  select(-para_time_point)

colnames(LA_control)[3] <- "control_clutches"

# rejoin infected and control dfs
LA_rel_vir_clutches <- full_join(LA_inf, LA_control)
LA_rel_vir_clutches$host_time_point <- as.factor(LA_rel_vir_clutches$host_time_point)

# make a separate file to use later for infected v. control comparison
LA_rel_vir_clutches_infvcontrol <- full_join(LA_inf, LA_control)
LA_rel_vir_clutches_infvcontrol$host_time_point <- as.factor(LA_rel_vir_clutches_infvcontrol$host_time_point)


LA_rel_vir_clutches <- as.data.frame(LA_rel_vir_clutches)

str(LA_rel_vir_clutches)

LA_rel_vir_2 <- full_join(LA_rel_vir_lifespan, LA_rel_vir_clutches)

str(LA_rel_vir_2)
```

```{r, calculating mean spore yield per clone}
str(LA_infected_pruned)

# get mean number clutches for each clone*treatment
stats_spores <- LA_infected_pruned %>%
  group_by(host_time_point, para_time_point, clone) %>%
  summarise(mn_spores = mean(spores_tot)) 

# for selecting contemporaneous pairs of hosts/parasites
stats_spores$host_time_point <- as.character(stats_spores$host_time_point)
stats_spores$para_time_point <- as.character(stats_spores$para_time_point)

LA_inf <- stats_spores %>% # select non-controls
  filter(para_time_point != "none") 

# join spore data with earlier calculations
LA_means <- full_join(LA_rel_vir_2, LA_inf)
LA_means$host_time_point <- as.factor(LA_means$host_time_point)

str(LA_means)
```

## Infected vs. uninfected comparison
Let's start by looking at lifespan of control/unexposed animals vs. infected animals. Based on what we know about the system, these may not be very different.

```{r, arranging data for inf v control comparison}
# Making a long data set with the uninfected & infected animals to compare lifespan
LA_rel_vir_lifespan_infvcontrol <- LA_rel_vir_lifespan_infvcontrol %>%
  rename(Infected = mn_lifespan)

LA_rel_vir_lifespan_infvcontrol <- LA_rel_vir_lifespan_infvcontrol %>%
  rename(Unexposed = control_lifespan)

LA_rel_vir_lifespan_infvcontrol_long <- LA_rel_vir_lifespan_infvcontrol %>%
  gather(infvcontrol, lifespan, Infected:Unexposed)

LA_rel_vir_lifespan_infvcontrol_long <- na.omit(LA_rel_vir_lifespan_infvcontrol_long)
  

# Now do the same for the reproduction data
LA_rel_vir_clutches_infvcontrol <- LA_rel_vir_clutches_infvcontrol %>%
  rename(Infected = mn_clutches)

LA_rel_vir_clutches_infvcontrol <- LA_rel_vir_clutches_infvcontrol %>%
  rename(Unexposed = control_clutches)

LA_rel_vir_clutches_infvcontrol_long <- LA_rel_vir_clutches_infvcontrol %>%
  gather(infvcontrol, clutches, Infected:Unexposed)

LA_rel_vir_clutches_infvcontrol_long <- na.omit(LA_rel_vir_clutches_infvcontrol_long)

```

```{r, plotting lifespan data for infvcontrol}
lifespaninfvcontrol <- ggplot(LA_rel_vir_lifespan_infvcontrol_long, 
                               aes(x=infvcontrol, y=lifespan)) + 
  geom_violin() + 
  geom_jitter(shape=16, position=position_jitter(width=0.3, height=0), alpha = 0.5, color = '#5c1a33') +
  xlab("Infected or unexposed") + 
  ylab("Mean lifespan \n(days)")  +
  theme_cowplot()

lifespaninfvcontrol

```



```{r, analysis of host lifespan for inf v control}
infvcontrollifespanmodel <- glm(lifespan ~ infvcontrol, 
             family = "quasipoisson", 
             data = LA_rel_vir_lifespan_infvcontrol_long)

plot(infvcontrollifespanmodel) 
summary(infvcontrollifespanmodel)

lifespansummary <- LA_rel_vir_lifespan_infvcontrol_long %>%
  group_by(infvcontrol) %>%
  summarise(mean = mean(lifespan))

lifespansummary

tapply(LA_rel_vir_lifespan_infvcontrol_long$lifespan, LA_rel_vir_lifespan_infvcontrol_long$infvcontrol, mean)
```


### Fecundity
Expectation: infected ones should have many fewer clutches.
```{r, plotting fecundity data for infected v control comparison}
# making a plot comparing exposed vs. unexposed
fecundityinfvcontrol <- ggplot(LA_rel_vir_clutches_infvcontrol_long, 
                               aes(x=infvcontrol, y=clutches)) + 
  geom_violin() + 
  geom_jitter(shape=16, position=position_jitter(width=0.3, height=0), alpha = 0.5, color = '#5c1a33') +
  xlab("Infected or unexposed") +ylab("Mean # clutches")  +
  theme_cowplot()

fecundityinfvcontrol
```

```{r, analysis of host repro for inf v control}
LA_rel_vir_clutches_infvcontrol_long$log_clutches <- log((LA_rel_vir_clutches_infvcontrol_long$clutches+1))

reprosummary <- LA_rel_vir_clutches_infvcontrol_long %>%
  group_by(infvcontrol) %>%
  summarise(mean = mean(clutches))

reprosummary

infvcontrolrepromodel <- glm(log_clutches ~ infvcontrol, 
             family = "gaussian", 
             data = LA_rel_vir_clutches_infvcontrol_long)

plot(infvcontrolrepromodel) 
summary(infvcontrolrepromodel)

tapply(LA_rel_vir_clutches_infvcontrol_long$clutches, LA_rel_vir_clutches_infvcontrol_long$infvcontrol, mean)


```

Let's arrange the lifespan & fecundity plots into a single figure:
```{r, single plot of inf v unexposed}
infvcontrolplot <- plot_grid(lifespaninfvcontrol, fecundityinfvcontrol, labels = "auto", ncol = 1, align = "v")
infvcontrolplot

ggsave(here("figures", "infvcontrolplot.jpg"), infvcontrolplot, units = "in", width = 5, height = 8, dpi = 300)

```

## Analysis of infection prevalence in infection assays
```{r, import dataset that will allow analysis of prevalence in infection assays}
# The dataset we were working with above only has infected animals or controls. Let's upload the dataset that will also include animals that were exposed but did not become infected
LA_full <- readr::read_csv(here("data/LA_full.csv")) 

# since we want to look at infectivity, let's get rid of the control/unexposed animals
LA_exposed <- LA_full %>%
  subset(para_time_point != "none")

# get rid of time 4 hosts, since we didn't have parasites from that time point
LA_exposed <- LA_exposed %>%
  subset(host_time_point != '4')

# filter out ones that are missing spore counts
LA_exposed <- LA_exposed %>%
  drop_na(spores_tot)

# Looking for clones that had very few animals in the experiment
LA_exposed_samplesizes <- LA_exposed %>%
  count(clone, para_time_point)

LA_exposed_samplesizes$remove <- ifelse(LA_exposed_samplesizes$n < 3, 'y', 'n')

# save
write.csv(LA_exposed_samplesizes, here("tables", "LA_exposed_samplesizes.csv"), row.names = F)

# remove clone-parasite combinations that had fewer than 3 individuals   
LA_exposed$remove <- ifelse(LA_exposed$clone == '31' |
           LA_exposed$clone == '224' |
           LA_exposed$clone == '312' & LA_exposed$para_time_point == '1' |
           LA_exposed$clone == '329' & LA_exposed$para_time_point == '1' |
           LA_exposed$clone == '335', 1, 0) 

LA_exposed_pruned <- LA_exposed %>%
  subset(remove == 0)


# Using spore yield to determine if it was infected; this is more sensitive than the 'infection' column, which was based on a quick visual assessment at death -- that inspection missed some that had very low spore densities
LA_exposed_pruned$infected <- ifelse(LA_exposed_pruned$spores_tot == 0, 0, 1)


# Calculate number infected and total sample sizes
LA_exposed_prevalence <- LA_exposed_pruned %>%
  group_by(clone, host_time_point, para_time_point) %>%
  summarise(total = n(), inf = sum(infected)) 
  
# Calculate proportion infected per clone
LA_exposed_prevalence$propinf <- LA_exposed_prevalence$inf/LA_exposed_prevalence$total

LA_exposed_prevalence$host_time_point <- as.character(LA_exposed_prevalence$host_time_point)

infprevplot <- ggplot(LA_exposed_prevalence, 
                     aes(x=host_time_point, y=propinf, fill=para_time_point)) + 
  geom_violin(position=position_dodge(1)) + 
  geom_jitter(shape=16, position=position_jitter(width=0.3, height=0), alpha = 0.5, show.legend = FALSE) +
  scale_fill_manual(values=c("#5445b1", "#749dae", "#f3c483")) + 
  scale_color_manual(values=c("#5445b1", "#749dae", "#f3c483")) +
  xlab("Host time point") +ylab("Proportion infected") +
  labs(color= "Parasite time point") + labs(fill="Parasite time point") +
  theme_cowplot()

infprevplot

infprevplota <- LA_exposed_prevalence %>%
  filter(host_time_point == para_time_point) %>%
  ggplot(aes(x=host_time_point, y=propinf, fill=para_time_point)) + 
  geom_violin(position=position_dodge(1), show.legend = FALSE) + 
  geom_jitter(shape=16, position=position_jitter(width=0.3, height=0), alpha = 0.5, show.legend = FALSE) +
  scale_fill_manual(values=c("#5445b1", "#749dae", "#f3c483")) + 
  scale_color_manual(values=c("#5445b1", "#749dae", "#f3c483")) +
  ylim(0,1.00) +
  xlab("Host time point") +ylab("Proportion infected") +
  labs(color= "Parasite time point") + labs(fill="Parasite time point") +
  theme_cowplot()

infprevplota

infprevplotb <- LA_exposed_prevalence %>%
  filter(host_time_point == 3) %>%
  ggplot(aes(x=para_time_point, y=propinf, fill=para_time_point)) + 
  geom_violin(position=position_dodge(1), show.legend = FALSE) + 
  geom_jitter(shape=16, position=position_jitter(width=0.3, height=0), alpha = 0.5, show.legend = FALSE) +
  scale_fill_manual(values=c("#C0C0C0", "#f3c483")) + 
  scale_color_manual(values=c("#C0C0C0", "#f3c483")) +
  ylim(0,1.00) +
  xlab("Parasite time point") +ylab("Proportion infected") +
  theme_cowplot()

infprevplotb

infprevplot2 <- plot_grid(infprevplota, infprevplotb, labels = "auto", ncol = 2, align = "v")
infprevplot2

ggsave(here("figures", "infprevplot.jpg"), infprevplot2, units = "in", width = 8, height = 5, dpi = 300)
```

```{r, analyze infection prevalence data}
LA_exposed_prevalence_sametime <- subset(LA_exposed_prevalence, host_time_point == para_time_point)

# Version without clone
prevmodel1<-glm(cbind(inf, total-inf) ~ para_time_point, family='binomial', data=LA_exposed_prevalence_sametime) 
summary(prevmodel1)

# Testing for overdispersion (using code from Ben Bolker via Michelle Fearon)
overdisp_fun <- function(model) {
  rdf <- df.residual(model)
  rp <- residuals(model,type="pearson")
  Pearson.chisq <- sum(rp^2)
  prat <- Pearson.chisq/rdf
  pval <- pchisq(Pearson.chisq, df=rdf, lower.tail=FALSE)
  c(chisq=Pearson.chisq,ratio=prat,rdf=rdf,p=pval)
}

overdisp_fun(prevmodel1)

# that shows overdispersion (p < 0.05)
# Adding in clone to try to deal with overdispersion
prevmodel1b<-glmer(cbind(inf, total-inf) ~ para_time_point + (1 | clone), family='binomial', data=LA_exposed_prevalence_sametime) 

overdisp_fun(prevmodel1b)
summary(prevmodel1b)

LA_exposed_prevalence_sametime$para_time_point <- as.factor(LA_exposed_prevalence_sametime$para_time_point)

# Testing for overall effect of parasite time point
drop1(prevmodel1b,test="Chisq")

# Now comparing the two sets of spores that were used to infect time 3 hosts
LA_exposed_prevalence_time3 <- subset(LA_exposed_prevalence, host_time_point == '3')

prevmodel2<-glm(cbind(inf, total-inf) ~ para_time_point, family='binomial', data=LA_exposed_prevalence_time3)

overdisp_fun(prevmodel2)

prevmodel2b<-glmer(cbind(inf, total-inf) ~ para_time_point + (1 | clone), family='binomial', data=LA_exposed_prevalence_time3) 

overdisp_fun(prevmodel2b)

summary(prevmodel2b)

# Getting chi-square to be parallel to initial analysis
drop1(prevmodel2b,test="Chisq")
```


## Parasite evolution comparison
Comparing hosts from time 3 that were exposed to parasites from time 1 vs. time 3

```{r, get dataset for just hosts from time 3}
LA_means_paraevol <- subset(LA_means, host_time_point == 3)
```

### Lifespan
```{r, lifespan for hosts from time 3}
lifespanparaevol <- ggplot(LA_means_paraevol, 
                     aes(x=para_time_point, y=mn_lifespan, fill=para_time_point)) + 
  geom_violin(show.legend = FALSE) + 
  geom_jitter(shape=16, position=position_jitter(width=0.3, height=0), alpha = 0.5, show.legend = FALSE) +
  scale_fill_manual(values=c("#5445b1", "#f3c483")) + 
  scale_color_manual(values=c("#5445b1", "#f3c483")) +
  xlab("Parasite time point") +ylab("Mean lifespan \n(days)") +
  labs(color= "Parasite time point") + labs(fill="Parasite time point") +
  theme_cowplot()


lifespanparaevol

```

```{r, analysis of host lifespan for parasite evolution comparison}
mod2a <- glm(mn_lifespan ~ para_time_point, 
             family = "quasipoisson", 
             data = LA_means_paraevol)

plot(mod2a) 
summary(mod2a)
```


### Reproduction
```{r, reproduction for hosts from time 3}
reproparaevol <- ggplot(LA_means_paraevol, 
                     aes(x=para_time_point, y=mn_clutches, fill=para_time_point)) + 
  geom_violin(show.legend = FALSE) + 
  geom_jitter(shape=16, position=position_jitter(width=0.3, height=0), alpha = 0.5, show.legend = FALSE) +
  scale_fill_manual(values=c("#5445b1", "#f3c483")) + 
  scale_color_manual(values=c("#5445b1", "#f3c483")) +
  xlab("Parasite time point") +ylab("Mean # clutches") +
  labs(color= "Parasite time point") + labs(fill="Parasite time point") +
  theme_cowplot()


reproparaevol

```

```{r, analysis of host reproduction for parasite evolution comparison}
LA_means_paraevol$log_mn_clutches <- log((LA_means_paraevol$mn_clutches+1))

mod2b <- glm(log_mn_clutches ~ para_time_point, 
             family = "quasipoisson", 
             data = LA_means_paraevol)

plot(mod2b) 
summary(mod2b)
```


### Spores
```{r, spore yield for hosts from time 3}
sporesparaevol <- ggplot(LA_means_paraevol, 
                     aes(x=para_time_point, y=mn_spores, fill=para_time_point)) + 
  geom_violin(show.legend = FALSE) + 
  geom_jitter(shape=16, position=position_jitter(width=0.3, height=0), alpha = 0.5, show.legend = FALSE) +
  scale_fill_manual(values=c("#5445b1", "#f3c483")) + 
  scale_color_manual(values=c("#5445b1", "#f3c483")) +
  xlab("Parasite time point") +ylab("Mean # spores \nper infected host") +
  labs(color= "Parasite time point") + labs(fill="Parasite time point") +
  scale_y_log10() +
  theme_cowplot()


sporesparaevol

```

```{r, analysis of spore yield for parasite evolution comparison}
LA_means_paraevol$log_mn_spores <- log(LA_means_paraevol$mn_spores)

mod2c <- glm(log_mn_spores ~ para_time_point, 
             family = "gaussian", 
             data = LA_means_paraevol)

plot(mod2c) 
summary(mod2c)
```

### Parasite growth rate
```{r, parasite growth rate for hosts from time 3}
LA_means_paraevol$paragrowth <- (LA_means_paraevol$mn_spores)/(LA_means_paraevol$mn_lifespan)

paragrowthparaevol <- ggplot(LA_means_paraevol, 
                     aes(x=para_time_point, y=paragrowth, fill=para_time_point)) + 
  geom_violin(show.legend = FALSE) + 
  geom_jitter(shape=16, position=position_jitter(width=0.3, height=0), alpha = 0.5, show.legend = FALSE) +
  scale_fill_manual(values=c("#5445b1", "#f3c483")) + 
  scale_color_manual(values=c("#5445b1", "#f3c483")) +
  xlab("Parasite time point") +ylab("Parasite growth rate \n(spores per day)") +
  labs(color= "Parasite time point") + labs(fill="Parasite time point") +
  theme_cowplot()


paragrowthparaevol

```

```{r, analysis of parasite growth rate for parasite evolution comparison}
mod2d <- glm(paragrowth ~ para_time_point, 
             family = "gaussian", 
             data = LA_means_paraevol)

plot(mod2d) 
summary(mod2d)
```

### Combined figure 
```{r, figure with plots for parasite evolution comparisons}
paraevolplot <- plot_grid(lifespanparaevol, reproparaevol, sporesparaevol, paragrowthparaevol, labels = "auto", ncol = 2, align = "v")
paraevolplot

ggsave(here("figures", "paraevolplot.jpg"), paraevolplot, units = "in", width = 10, height = 7, dpi = 300)
```

```{r, plots of parasite growth rate}
clutchessporesplot <- ggplot(LA_means_paraevol, aes(x = mn_clutches, y = mn_spores)) +
  geom_jitter(shape = 16, position=position_jitter(width=0.3, height=0.3), alpha = 0.8, aes(color=para_time_point)) +
  geom_smooth(method='lm') +
  theme_cowplot()

clutchessporesplot 

lifespansporesplot <- ggplot(LA_means_paraevol, aes(x = mn_lifespan, y = mn_spores)) +
  geom_jitter(shape = 16, position=position_jitter(width=0.3, height=0.3), alpha = 0.8, aes(color=para_time_point)) +
  geom_smooth(method='lm') +
  theme_cowplot()

lifespansporesplot 

clutchessporegrowthplot <- ggplot(LA_means_paraevol, aes(x = mn_clutches, y = paragrowth)) +
  geom_jitter(shape = 16, position=position_jitter(width=0.3, height=0.3), alpha = 0.8, aes(color=para_time_point)) +
  geom_smooth(method='lm') +
  theme_cowplot()

clutchessporegrowthplot 

lifespansporegrowthplot <- ggplot(LA_means_paraevol, aes(x = mn_lifespan, y = paragrowth)) +
  geom_jitter(shape = 16, position=position_jitter(width=0.3, height=0.3), alpha = 0.8, aes(color=para_time_point)) +
  geom_smooth(method='lm') +
  theme_cowplot()

lifespansporegrowthplot 


time3sporeplot <- plot_grid(clutchessporesplot, lifespansporesplot, clutchessporegrowthplot, lifespansporegrowthplot, labels = "auto", ncol = 2, align = "v")
time3sporeplot

ggsave(here("figures", "time3sporeplot.jpg"), time3sporeplot, units = "in", width = 10, height = 7, dpi = 300)
```


```{r, correlations with parasite growth rate}
cor.test(LA_means_paraevol$log_mn_clutches, LA_means_paraevol$log_mn_spores)

cor.test(LA_means_paraevol$log_mn_clutches, LA_means_paraevol$mn_lifespan)

cor.test(LA_means_paraevol$mn_lifespan, LA_means_paraevol$log_mn_spores)

cor.test(LA_means_paraevol$log_mn_clutches, LA_means_paraevol$mn_spores)

cor.test(LA_means_paraevol$mn_lifespan, LA_means_paraevol$log_mn_spores)

cor.test(LA_means_paraevol$log_mn_clutches, LA_means_paraevol$paragrowth)

cor.test(LA_means_paraevol$mn_lifespan, LA_means_paraevol$paragrowth)

```

